Update of the ozone dry deposition in the OsloCTM3 v1.0

### 2018-11-20 DryDep.tex
Finished section about zonal distribution, started with seasonal cycles. Had to update a few plots. No spell-check so far...

The phenology of a plant typically describes its life-cycle through out a year, e.g., in mid latitudes and for deciduous species, it starts with the emergence of leafs in spring and ends in fall. In the EMEP scheme, phenology is parameterized with respect to the start of the greening season (SGS) and its end (EGS). Details about our treatment of these are given in Section~\ref{subsubsec:greening}. In summary, our adaption of the $f_\text{phen}$ parameterization reads as follows:
%
\begin{equation}
  f_\text{phen} =
  \begin{cases}
    \emph{if} \quad \text{GDAY = 0}  & \quad 0 \\
    \emph{if} \quad \text{GLEN $\ge$ 365} & \quad 1 \quad \emph{(explicitly excluding Tropics)} \\
    \emph{else}  & \quad 
    \begin{cases}
      \emph{if} \quad \text{GDAY} \le \phi_\text{AS}  & \quad \phi_a \\
      \emph{if} \quad \text{GDAY} \le \phi_\text{AS} + \phi_e  & \quad \phi_b+(\phi_c-\phi_b)\cdot(\text{GDAY}-\phi_\text{AS})/\phi_e\\
      \emph{if} \quad \text{GDAY} \le \text{GLEN} - \phi_\text{AE} - \phi_f  & \quad \phi_c\\
      \emph{if} \quad \text{GDAY} \le \text{GLEN} - \phi_\text{AE}  & \quad \phi_d+(\phi_c-\phi_d)\cdot(\text{GLEN}-\phi_\text{AE}-\text{GDAY})/\phi_f\\
      \emph{else} & \quad \phi_d
    \end{cases}
  \end{cases}
  \label{eq:fphen}
\end{equation}
%
Herein, we use the SGS and EGS derived parameters day of greening season (GDAY), the time elapsed starting at the SGS, and the total length of the greening season (GLEN), the time span between EGS and SGS. The parameters $\phi_a$, $\phi_b$, $\phi_c$, and $\phi_d$ define start or end points in the five phases of phenology in the EMEP scheme, while $\phi_e$, $\phi_f$, $\phi_\text{AS}$, and $\phi_\text{AE}$ control the temporal timing (Fig.~\ref{fig:phenology}). If GLEN is zero we are most likely in Arctic regions and there is no vegetation anyway, therefore $f_\text{phen}=0$. Before the start of the growing season ($\text{GDAY}=0$) $f_\text{phen}=0$. Since the EMEP scheme is tuned to northern hemisphere (NH) mid latitudes, this phenology does not apply to the tropics, after evaluation of the model results (Section~\ref{subsec:sens}), we decided to set $f_\text{phen}=1$ if GLEN is greater or equal to 365 which is the case in the tropics.
%
\begin{figure}[t]
  \includegraphics[width=8.3cm]{pictures/phenology.pdf}
  \caption{Sketch of the five different phases described in $f_\text{phen}$.}
  \label{fig:phenology}
\end{figure}

@@

The resulting PPFD fields are still accumulated over a time period of three hours and should be divided by $3$. A \href{https://confluence.ecmwf.int/display/CKB/ERA-Interim\%3A+surface+photosynthetically+active+radiation+\%28surface+PAR\%29+values+are+too+low}{known issue} in the OpenIFS (cycles $\le$ c41r2) causes surface PAR values to be about $30\,\unit{\%}$ below observations. To counter this, we decided to refrain from the division at this stage, but need to bare this in mind for later OpenIFS versions.

@@

In this section, we present results from a manifold of OsloCTM3 model integrations testing different parameters of the EMEP scheme. We focus on changes in ozone related to the stomatal conductance parameterization, e.g., total dry deposition $\sum\chem{O_3^\text{DD}}$, dry deposition velocities $v^\chem{O_3}_\text{DD}$, and surface concentration $[\chem{O_3}](z_0)$ or surface burden $\chem{O_3}(z_0)$ in units of \unit{kg}, respectively. We evaluate our results with respect to the multi-model evaluation of ozone dry deposition by \citet{ACP:Hardacre2015} (Section~\ref{subsec:model}) and observations (Section~\ref{subsec:obs}). The OsloCTM3 is driven by meteorological input fields from \href{https://www.ecmwf.int/en/forecasts/documentation-and-support/evolution-ifs/cycle-38r1-summary-changes}{ECMWF -- OpenIFS version cy38r1nc4}. CEDS historical emission inventory is used for anthropogenic emissions, while biomass burning is covered in daily resolution by NASA's Global Fire Emissions Database, Version4 (GFEDv4). Biogenic emissions are taken from MEGAN output. In the following, we will present the various model sensitivity studies.

@@

As presented in Section~\ref{sec:model_des}, the OsloCTM3 dry deposition scheme has undergone a major revision. Due to the previously generalized implementation of the Wesely scheme with respect to dry deposition resistances and surface types, it is not possible to fully disentangle the effects of every single change. Therefor, the conducted sensitivity studies (9 in total) mainly assess the parameter space of the stomatal conductance within the EMEP scheme. A reference simulation featuring the Wesely scheme has been done and will be referred to as \emph{Wesely\_type}, indicating that other implementations of the original work by \citet{AE:Wesely1989} may exist in other models. For all model integrations, the meteorological reference year is 2005. This choice only affects the comparison with data and multi-model studies that either perform analysis on decadal averages or differing years. After finishing the work on the integration of the EMEP parameterization and initial testing, \emph{EMEP\_full} was supposed to be the baseline simulation for our sensitivity studies, \emph{EMEP\_offLight} and \emph{EMEP\_offPhen}. As indicated by the names, these scenarios are extreme cases, setting $f_\text{light}$ and $f_\text{phen}$ to 1, respectively. During analysis of the results, we found, that the SMOKE-BEIS parameterization of the greening season did not extend to the boreal and subarctic regions. Especially in Europe, this basically results in the suppression of canopy resistance in these regions. The following sensitivity study therefor comprises two different versions of pre-processed greening season (refer to Section~\ref{subsec:pre-pro} for details), mid and high latitudes in the northern hemisphere only (\emph{EMEP\_ppgs}) and for both, northern and southern hemisphere mid and high latitudes (\emph{EMEP\_ppgssh}). Building on the latter, we want to confirm the importance of the $v^\chem{O_3}_\text{ice/snow}$ update \citep{ACP:Helmig2007} in the Subarctic and Arctics, within the framework of the OsloCTM3 (\emph{EMEP\_ppgssh\_ice}). Accidentally, we have used forcings for the year 2014 instead of 2005. Though, we take this as opportunity to assess the impact on ozone regarding differing emissions (\emph{EMEP\_ppgs\_2005}). As mentioned previously, the simulations initially ran with the uppermost soil water level (SWVL1) in the implementation of $f_\text{SW}$. A final sensitivity study was conducted changing to soil water level at depths deeper than $1\,\unit{m}$ (\emph{EMEP\_SWVL4}).

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Comparison with multi-model evaluation}
\label{subsec:model}
%
In our evaluation of results, we closly follow suggestions by \citet{ACP:Hardacre2015}. For the purpose of comparison with the multi-model mean of the therein participating Task Force on Hemispheric Transport of Air Pollution (TF HTAP) models, we also have re-grided our data to a horizontal resolution of $3^\circ\times 3^\circ$. First (Section~\ref{subsubsec:zonal}), we look at zonal ditributions of $[\chem{O_3}](z_0)$, $v^\chem{O_3}_\text{DD}$, and $\sum\chem{O_3^\text{DD}}$ for all our senstivity simulations. In Section~\ref{subsubsec:seasons}, we study seasonal cycles of hemispheric ozone as well as nine land surface types which have been retrospectively seperated. From this, we estimate the total annual ozone dry deposition onto ocean, ice, and land surfaces and compare with results from \citet{ACP:Luhar2017}.
%
\subsubsection{Zonal distribution}
\label{subsubsec:zonal}
%
The annual zonal average with respect to surface ozone concentration (Fig.~\ref{fig:mmm_drydep}a) displays, in consistancy with Fig.~\ref{fig:diff_maps}a), a general, global increase by up to $10\,\unit{ppb}$ which is largest between $(25-50)\,\unit{^\circ N}$ including the major deserts. In the deep Tropics, the increase is smallest ($\Delta[\chem{O_3}]\leq 5\,\unit{ppb}$). On average, we find that the EMEP scheme further intensifies the strong asymmetry between northern and southern hemisphere. Among the senistivity studies, there is only a low variance. Most remarkable and as expected due to the strong decrease in prescribed dry deposition velocity over ice and snow, \emph{EMEP\_ppgssh\_ice} displays almost a doubling of surface ozone in the high Arctics compared to \emph{Wesely\_type} but affects ozone concentrations already at about $50\,\unit{^\circ}$ in both hemispheres. Neglecting any dependence on solar radiation in the stomatal conductance formulation (\emph{EMEP\_offLight}) generally decreases the ozone concentration by about $2\,\unit{ppb}$, while taking deeper soil water content into consideration (\emph{EMEP\_SWVL4}) increases $[\chem{O_3}](z_0)$ slightly. Rather surprisingly, switching of phenology completely (\emph{EMEP\_offPhen}), doesn't make much difference. The scenario of differing emissions (2005 in comparison to 2014 or more specifically \emph{EMEP\_ppgs\_2005} compared to \emph{EMEP\_ppgs}), yields higher ozone concentrations in the northern hemisphere in 2005 in accordance to a reduction in sulfur and \chem{NO_x} emissions in south east Asia in later years. An opposite tendency is seen for latitudes south of $30\,\unit{^\circ N}$, propbably owing to the industrial development of countries in these regions.

In Fig.~\ref{fig:mmm_drydep}b), we show the corresponding average dry deposition velocities, as reconstructed from the ration between surface concentration and total dry deposition since we have no output of these. $v^\chem{O_3}_\text{DD}$ in the EMEP scheme are much smaller then in the Wesely scheme even smaller compared to the multi-model-mean for latitudes between $50\,\unit{^\circ S}$ and $70\,\unit{^\circ N}$. In the remaining latitude bands, except for EMEP\_ppgssh\_ice, the EMEP scheme is constantly above the multi-model-mean. This indicates, that with respect to other models, the updated dry deposition velocity above ice and snow should be considered as new standard.
  \includegraphics[width=8.3cm]{pictures/final-vmr_surfaceozone}\\
  \includegraphics[width=8.3cm]{pictures/final-total_ozone_drydepvelo}\\
  \includegraphics[width=8.3cm]{pictures/final-total_ozone_drydep}
  \caption{Comparison of the manifold of OsloCTM3 integrations with respect to (a) Surface ozone concentrations, (b) Annual average ozone dry deposition velocity, (c) Total annual ozone dry deposition. The different colors indicate sets of simulation with similar baselines. The multi-model mean from the evaluation of HTAP models by \citet{ACP:Hardacre2015} is shown as a reference (where available).}
  \label{fig:mmm_drydep}
  \includegraphics[width=8.3cm]{pictures/final-total_ozone_drydep_monthly}\\
  \includegraphics[width=8.3cm]{pictures/final-total_ozone_drydepvelo_monthly}
  \caption{Comparison of monthly zonal distributions to results from \citep{ACP:Hardacre2015}. (a) Total amount of ozone removed from the atmosphere; (b) average ozone dry deposition velocities.}
  \label{fig:mmm_drydep_season}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Average seasonal cycles}
\label{subsubsec:seasons}

@@

  \caption{Seasonal cycle of total amount of ozone removed from the atmosphere through dry deposition separated into northern hemisphere (NH), tropics (TR), and southern hemisphere (SH). Comparison with results from \citet{ACP:Hardacre2015}.}

>>>>

The phenology of a plant typically describes its life-cycle through out a year, e.g., in mid latitudes and for deciduous species, it starts with the emergence of leafs in spring and ends in fall. In the EMEP scheme, phenology is parameterized with respect to the start of the greening season (SGS) and its end (EGS). Details about our treatment of these are given in Section~\ref{subsubsec:greening}. In summary, our adaption of the $f_\text{phen}$ parameterization reads as follows:
%
\begin{equation}
  f_\text{phen} =
  \begin{cases}
    \emph{if} \quad \text{GDAY = 0}  & \quad 0 \\
    \emph{if} \quad \text{GLEN $\ge$ 365} & \quad 1 \quad \emph{(explicitly excluding tropics)} \\
    \emph{else}  & \quad 
    \begin{cases}
      \emph{if} \quad \text{GDAY} \le \phi_\text{AS}  & \quad \phi_a \\
      \emph{if} \quad \text{GDAY} \le \phi_\text{AS} + \phi_e  & \quad \phi_b+(\phi_c-\phi_b)\cdot(\text{GDAY}-\phi_\text{AS})/\phi_e\\
      \emph{if} \quad \text{GDAY} \le \text{GLEN} - \phi_\text{AE} - \phi_f  & \quad \phi_c\\
      \emph{if} \quad \text{GDAY} \le \text{GLEN} - \phi_\text{AE}  & \quad \phi_d+(\phi_c-\phi_d)\cdot(\text{GLEN}-\phi_\text{AE}-\text{GDAY})/\phi_f\\
      \emph{else} & \quad \phi_d
    \end{cases}
  \end{cases}
  \label{eq:fphen}
\end{equation}
%
Herein, we use the SGS and EGS derived parameters day of greening season (GDAY), the time elapsed starting at the SGS, and the total length of the greening season (GLEN), the time span between EGS and SGS. The parameters $\phi_a$, $\phi_b$, $\phi_c$, and $\phi_d$ define start or end points in the five phases of phenology in the EMEP scheme, while $\phi_e$, $\phi_f$, $\phi_\text{AS}$, and $\phi_\text{AE}$ control the temporal timing (Fig.~\ref{fig:phenology}). If GLEN is zero we are most likely in Arctic regions and there is no vegetation anyway, therefore $f_\text{phen}=0$. Before the start of the growing season ($\text{GDAY}=0$) $f_\text{phen}=0$. Since the EMEP scheme is tuned to northern hemisphere (NH) mid latitudes, this phenology does not apply to the tropics, after evaluation of the model results (Section~\ref{subsec:sens}), we decided to set $f_\text{phen}=1$ if GLEN is greater or equal to 365 which is the case in the tropics.
%
\begin{figure}[t]
  \includegraphics[width=8.3cm]{pictures/phenology.pdf}
  \caption{Sketch of the five different phases described in $f_\text{phen}$.}
  \label{fig:phenology}
\end{figure}

@@

The resulting PPFD fields are still accumulated over a time period of three hours and should be divided by $3$. A \href{https://confluence.ecmwf.int/display/CKB/ERA-Interim\%3A+surface+photosynthetically+active+radiation+\%28surface+PAR\%29+values+are+too+low}{known issue} in the OpenIFS (cycles $\le$ c41r2) causes surface PAR values to be about $30\,\unit{\%}$ below observations. To counter this, we decided to refrain from the division at this stage, but need to bare this in mind for later OpenIFS cycles.

@@

In this section, we present results from a manifold of OsloCTM3 model integrations testing different parameters of the EMEP scheme. We focus on changes in ozone related to the stomatal conductance parameterization, e.g., total dry deposition $\sum\chem{O_3^\text{DD}}$, dry deposition velocities $v^\chem{O_3}_\text{DD}$, and surface concentration $[\chem{O_3}](z_0)$ or surface burden $\chem{O_3}(z_0)$ in units of \unit{kg}, respectively. We evaluate our results with respect to the multi-model evaluation of ozone dry deposition by \citet{ACP:Hardacre2015} (Section~\ref{subsec:model}) and observations (Section~\ref{subsec:obs}). The OsloCTM3 is driven by meteorological input fields from \href{https://www.ecmwf.int/en/forecasts/documentation-and-support/evolution-ifs/cycle-38r1-summary-changes}{ECMWF -- OpenIFS cy38r1nc4}. CEDS historical emission inventory is used for anthropogenic emissions, while biomass burning is covered in daily resolution by NASA's Global Fire Emissions Database, Version4 (GFEDv4). Biogenic emissions are taken from MEGAN output. In the following, we will present the various model sensitivity studies.

@@

Due to the divergence between the implementation of the EMEP scheme and the previous Wesely scheme with respect to dry deposition velocities and surface types, it is not possible to fully disentangle and trace back differing results to a single change. Therefor, we conducted in total nine sensitivity studies that mainly assess the parameter space of the stomatal conductance within the EMEP scheme. A reference simulation featuring the Wesely scheme has been done and will be referred to as \emph{Wesely\_type}, indicating that other implementations of the original work by \citet{AE:Wesely1989} may exist in other models. For all model integrations, the meteorological reference year is 2005. This choice only affects the comparison with data and multi-model studies that either perform analysis on decadal averages or differing years. After finishing the work on the integration of the EMEP parameterization and initial testing, \emph{EMEP\_full} was supposed to be the baseline simulation for our sensitivity studies, \emph{EMEP\_offLight} and \emph{EMEP\_offPhen}. As indicated by the names, these scenarios are extreme cases, setting $f_\text{light}$ and $f_\text{phen}$ to 1, respectively. During analysis of the results, we found, that the SMOKE-BEIS parameterization of the greening season did not extend to the boreal and subarctic regions. Especially in Europe, this basically results in the suppression of canopy resistance in northern Scandinavia. The following sensitivity study therefor comprises two different versions of pre-processed greening season (refer to Section~\ref{subsec:pre-pro} for details), mid and high latitudes in the northern hemisphere only (\emph{EMEP\_ppgs}) and for both, northern and southern hemisphere mid and high latitudes (\emph{EMEP\_ppgssh}). Building on the latter, we want to confirm the importance of the $v^\chem{O_3}_\text{ice/snow}$ update \citep{ACP:Helmig2007} in the Subarctic and Arctics, within the framework of the OsloCTM3 (\emph{EMEP\_ppgssh\_ice}). Accidentally, we have used forcings for the year 2014 instead of 2005. Though, we take this as opportunity to assess the impact on ozone regarding differing emissions (\emph{EMEP\_ppgs\_2005}). As mentioned previously, the simulations initially ran with the uppermost soil water level (SWVL1) in the implementation of $f_\text{SW}$. A final sensitivity study was conducted changing to soil water level at depths deeper than $1\,\unit{m}$ (\emph{EMEP\_SWVL4}).

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Comparison with multi-model evaluation}
\label{subsec:model}
%
In our evaluation of results, we closly follow suggestions by \citet{ACP:Hardacre2015}. For the purpose of comparison with the multi-model mean of the therein participating Task Force on Hemispheric Transport of Air Pollution (TF HTAP) models, we also have re-grided our data to a horizontal resolution of $3^\circ\times 3^\circ$. In Section~\ref{subsubsec:zonal}, we look at zonal ditributions of $[\chem{O_3}](z_0)$, $v^\chem{O_3}_\text{DD}$, and $\sum\chem{O_3^\text{DD}}$ for all our senstivity simulations and study seasonal cycles of hemispheric ozone as well as nine land surface types which have been retrospectively seperated (Section~\ref{subsubsec:seasons}). From this, we estimate the total annual ozone dry deposition onto ocean, ice, and land surfaces and compare with results from \citet{ACP:Luhar2017}.
%
\subsubsection{Zonal distribution}
\label{subsubsec:zonal}
%
The annual zonal average with respect to surface ozone concentration (Fig.~\ref{fig:mmm_drydep}a) displays, in consistancy with Fig.~\ref{fig:diff_maps}a), a general, global increase of ozone by up to $10\,\unit{ppb}$ which is largest between $(25-50)\,\unit{^\circ N}$ including the major deserts. In the deep tropics ($5\,\unit{^\circ S}-5\,\unit{^\circ N}$), the increase is smallest ($\Delta[\chem{O_3}]\leq 5\,\unit{ppb}$). We find that the EMEP scheme further intensifies the strong asymmetry between northern and southern hemisphere as a consequence of the distribution of the continental land masses and vegetation thereon. Among the sensitivity studies, there is only a low variance. Most remarkable and as expected due to the strong decrease in prescribed dry deposition velocity over ice and snow, \emph{EMEP\_ppgssh\_ice} displays almost a doubling of surface ozone in the high Arctics compared to \emph{Wesely\_type} but affects ozone concentrations already at about $50\,\unit{^\circ}$ in both hemispheres. Neglecting any dependence on solar radiation in the stomatal conductance formulation (\emph{EMEP\_offLight}) -- or in other words allowing photosynthesis 24/7 -- generally decreases the ozone concentration by about $2\,\unit{ppb}$, while choosing soil water at deeper levels than the surface layer (\emph{EMEP\_SWVL4}) increases $[\chem{O_3}](z_0)$ slightly. Rather surprisingly, switching of phenology completely (\emph{EMEP\_offPhen}), doesn't make much difference. The scenario of differing emissions (2005 in comparison to 2014 or more specifically \emph{EMEP\_ppgs\_2005} compared to \emph{EMEP\_ppgs}), yields higher ozone concentrations in the northern hemisphere in 2005 in accordance to a reduction in sulfur and \chem{NO_x} emissions in south east Asia in later years. An opposite tendency is seen for latitudes south of $30\,\unit{^\circ N}$, propbably owing to the industrial development of countries in these regions.

In Fig.~\ref{fig:mmm_drydep}b), we show the corresponding average dry deposition velocities, as reconstructed from the ratio between surface concentration and total dry deposition since we have no direct output of these. The $v^\chem{O_3}_\text{DD}$ in the EMEP scheme are much smaller then in the Wesely scheme and about a factor of $3$ smaller compared to the multi-model-mean for latitudes between $50\,\unit{^\circ S}$ and $70\,\unit{^\circ N}$. In the remaining latitude bands, except for \emph{EMEP\_ppgssh\_ice}, all realisations of the EMEP scheme are constantly above the multi-model-mean. This indicates, that with respect to other models, the updated dry deposition velocity above ice and snow should be considered as new standard for the OsloCTM3. The dry deposition velocities are of cause independent of the emission scenario, but display a strong dependence on $f_\text{light}$ and $f_\text{phen}$. The closer to $1$ these factores are the higher do the dry deposition velocities become (compare \emph{EMEP\_offLight}, \emph{EMEP\_offPhen}, and \emph{EMEP\_ppgssh}). That highlights the importance of a proper PAR from meteorological input in CTMs or radiative transfer in PAR in global climate models (GCM) for canopy conductance computation. Albeit, the total numbers are smaller, the shape of the zonal average dry deposition velocities from the EMEP scheme and the multi-model-mean are similar (see figure in Supplement~S.6.1). The biggest exception is the zonal band $(12-30)\,\unit{^\circ S}$, which coincide with the location of Australia and its large desert regions.

The annual total ozone dry deposition is shown in Fig.~\ref{fig:mmm_drydep}b). In accordance to the priviously decribed features, we observe a reduction of the global total ozone dry deposition to almost half of \emph{Wesely\_type} in any of the sensitivity studies. The occurence of this reduction in the zonal bands where the major deserts are located may point to a important change in $v_\text{desert}^\chem{O_3}$. Consulting the old parameter file, we find $v_\text{desert}^\chem{O_3} \equiv v_\text{tundra}^\chem{O_3} = 0.26\,\unit{cm\,s^{-1}}$, while in the EMEP scheme $v_\text{desert}^\chem{O_3}=0.05\,\unit{cm\,s^{-1}}$. Similarly, dry deposition velocities over ice and snow and ocean have been even higher in the Wesely scheme than in the original EMEP scheme ($v_\text{ice/snow}^\chem{O_3}\equiv v_\text{water}^\chem{O_3}=0.07\,\unit{cm\,s^{-1}}$). These differences in baseline dry deposition velocities over huge parts of the unvegetated surface of the Earth (see Fig.~\ref{fig:pft_landsurface}) amount to most of the quantitative change we see between the Wesely and the EMEP scheme. We further elaborate on this in the following section.

\begin{figure}[t]
  \includegraphics[width=8.3cm]{pictures/final-vmr_surfaceozone}\\
  \includegraphics[width=8.3cm]{pictures/final-total_ozone_drydepvelo}\\
  \includegraphics[width=8.3cm]{pictures/final-total_ozone_drydep}
  \caption{Comparison of the manifold of OsloCTM3 integrations with respect to (a) Surface ozone concentrations, (b) Annual average ozone dry deposition velocity, (c) Total annual ozone dry deposition. The different colors indicate sets of simulation with similar baselines. The multi-model mean from the evaluation of HTAP models by \citet{ACP:Hardacre2015} is shown as a reference (where available).}
  \label{fig:mmm_drydep}
\end{figure}
%
In Fig.~\ref{fig:mmm_drydep_season}a), the average zonal ozone dry deposition is shown seperated by month. Where available, we have added the multi-model-mean given by \citet{ACP:Hardacre2015} as reference. There are two major phases apparent, NH and SH greening season. Former is reflected in a pronounced peak of $v_\text{DD}^\chem{O_3}$ in the northern mid latitudes, while the absence of this peak can be attributed to the latter in addition to a southward shift of the tropical peak dry deposition velocity. In the Wesely scheme, NH midlatitude peak velocities appear in June compared to July in the EMEP scheme, indicating that the prescribed seasonal cycles differ. The corresponding total monthly ozone dry deposition is shown in Fig.~\ref{fig:mmm_drydep_season}b). In general, the seasonal patters are quite similar in the Wesely scheme and the EMEP scheme, displaying a strong symmetry around $10\,\unit{^\circ N}$ in January/February and November/December, respectively. What differs most is the molding and intensitity of the NH peak dry deposition.Both schemes reach the maximum in June/July but the peak is much more differenciated in March already in the Wesely scheme. Similarly, the SH tropical peak dry deposition is reached in August/September but sustained longer into October in the Wesely scheme. Since we have not conducted any simulation with a meteorological year other than 2005, we cannot elaborate on whether this is a special feature of our chosen year or not.
%
\begin{figure}[t]
  \includegraphics[width=8.3cm]{pictures/final-total_ozone_drydepvelo_monthly}\\
  \includegraphics[width=8.3cm]{pictures/final-total_ozone_drydep_monthly}
  \caption{Comparison of the manifold of OsloCTM3 integrations with respect to (a) Zonal average ozone dry deposition velocities; (b) Total annual amount of ozone removed from the atmosphere via dry deposition. The multi-model mean from the evaluation of HTAP models by \citet{ACP:Hardacre2015} is shown as a reference (where available).}
  \label{fig:mmm_drydep_season}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Average seasonal cycles}
\label{subsubsec:seasons}
%
To disentangle the contributions of different regions to the global ozone budget further, we will briefly look at different projections of seasonal cycles.

In Fig.~\ref{fig:mmm_drydep_hem}, the total annual ozone dry deposition seperated into mid and high latitudes in the northern hemisphere ($(30-90)\,\unit{^\circ N}$), the tropics and subtropics ($30\,\unit{^\circ S}-30\,\unit{^\circ N}$), and the mid and high latitudes in the southern hemisphere ($(30-90)\,\unit{^\circ S}$) are shown. As reference, we have again added the multi-model-mean by \citet{ACP:Hardacre2015}. As expected, the NH mid and high latitudes display a strongly pronounced seasonal cycle, while it is less pronounced in the tropics and in the SH. The highest ozone dry deposition is found in the tropics and amounts on average to the  peak levelof dry deposition in the NH for the multi-model-mean \citep{ACP:Hardacre2015} and EMEP scheme. The average tropical ozone dry deposition in the Wesely scheme diverges by about $5\,\unit{Tg}$ in comparison to its NH maximum. The seasonal cycle in the EMEP scheme appears to be shifted towards later in the year in the NH. While the general seasonal cycle in the tropics and subtropics is not changing, we find the strongest sensitivity to the verious realisations of the EMEP scheme there, with \emph{EMEP\_offLight} and \emph{EMEP\_ppgs\_2005} the highest and the lowest, respectively. This indicates that surface ozone in the tropics may be very sensitive to changes to emissions and changes in longwave radiative transfer (e.g., cloud cover, aerosol).

@@

  \caption{Seasonal cycle of total annual amount of ozone removed from the atmosphere through dry deposition separated into northern hemisphere (NH), tropics (TR), and southern hemisphere (SH). The multi-model mean from the evaluation of HTAP models by \citet{ACP:Hardacre2015} is shown as a reference.}

### 2018-11-20 DryDep_supplement.tex
Finished section about zonal distribution, started with seasonal cycles. Had to update a few plots. No spell-check so far...

\subsection*{S.6 Average ozone dry deposition velocity seperated by land surface type (southern hemisphere)}
\appendixfigures
\begin{center}
  \includegraphics[height=0.4\textheight, clip, trim={0.7cm 0.7cm 0.7cm 0.55cm}]{pictures/final-total_ozone_drydepvelo_monthly_pft_sh}
\end{center}
\end{document}

>>>>

\subsection*{S.6 }
\subsubsection*{S.6.1 Normed zonal average ozone dry deposition velocity}
\appendixfigures
\begin{center}
  \includegraphics[height=0.4\textheight, clip, trim={0.7cm 0.7cm 0.7cm 0.55cm}]{pictures/final-norm_total_ozone_drydepvelo}
\end{center}

\subsubsection*{S.6.2 Average ozone dry deposition velocity seperated by land surface type (southern hemisphere)}
\appendixfigures
\begin{center}
  \includegraphics[height=0.4\textheight, clip, trim={0.7cm 0.7cm 0.7cm 0.55cm}]{pictures/final-total_ozone_drydepvelo_monthly_pft_sh}
\end{center}
\end{document}
